CF-SuGAR Power Calculations

### 2025-06-13 Christine Chan/CF-SuGAR/power_calculations.qmd
CF-SuGAR FEV1 power calculations

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(childsds)
library(rspiro)
library(survival)
library(gtsummary)
library(ggstats)
library(lmerTest)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Christine Chan",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

>>>>

```{r setup}
#| include: false
library(Hmisc)
library(tidyverse)
library(childsds)
library(rspiro)
library(survival)
library(gtsummary)
library(ggstats)
library(lmerTest)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

# Change in FEV1

```{r}
encounters <- read.csv("C:/Users/vigerst/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry/Data_Raw/CFF19_encountersMerged_Del1.csv")
# Exclude T1D and T2D
encounters <- encounters %>%
  filter(!cfrd_status %in% c(4, 5))
# Find each person's first year on modulators
encounters$Vx770[is.na(encounters$Vx770)] <- 0
encounters$Vx809comb[is.na(encounters$Vx809comb)] <- 0
encounters <- encounters %>%
  arrange(eDWID, reviewyear, encounternum) %>%
  group_by(eDWID) %>%
  mutate(
    modulator = factor(Vx770 == 1 | Vx809comb == 1,
      levels = c(F, T),
      labels = c("Not on Modulator", "On Modulator")
    ),
    first_modulator =
      first(reviewyear[modulator == "On Modulator"], na_rm = T)
  )
# Limit to two years prior and one year post-modulator
encounters <- encounters %>%
  filter(
    !is.na(first_modulator),
    reviewyear >= first_modulator - 2 & reviewyear <= first_modulator + 1
  )
# Limit dataset to people who were seen regularly but not too often
# (likely hospitalized), make a new encounter number
year_picker <- encounters %>%
  group_by(eDWID, reviewyear) %>%
  summarise(n = n()) %>%
  filter(n >= 3 & n <= 10) %>%
  ungroup() %>%
  mutate(id = paste(eDWID, reviewyear, sep = "_"))
encounters <- encounters %>%
  filter(paste(eDWID, reviewyear, sep = "_") %in% year_picker$id) %>%
  group_by(eDWID) %>%
  mutate(visit = row_number()) %>%
  select(
    eDWID, reviewyear, encounternum, visit, modulator,
    GLI_FEV1_pct_predicted
  )
# Plot
# ggplot(encounters, aes(x = visit, y = GLI_FEV1_pct_predicted, group = eDWID)) +
#   geom_line(alpha = 0.5)
# ggplot(encounters, aes(x = visit, y = GLI_FEV1_pct_predicted, group = eDWID)) +
#   geom_point(alpha = 0.1) +
#   geom_smooth(aes(color = modulator, group = modulator))
# Write for SAS models
write.csv(encounters, file = "C:/Users/vigerst/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF-SuGAR/Background/Power Calculations/Data/encounters.csv",row.names = F,na = "")
```
